Manuscript 1 code

### 2024-09-13 manuscript_1_code.Rmd
Added in 2 a-d for site and quad richness and evenness by year :cow:

```{r figure 1, fig.height = 10, fig.width=14, warning=FALSE}
(figure_1 <- ggplot() +
  geom_point(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
  geom_errorbar(data = figure_1_data, aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group),
                width = 0, size = 1.3) +
   geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 1) +
  facet_wrap(~ habitat) +
  annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
  scale_colour_manual(values = functional_group_colours) +
  scale_fill_manual(values = functional_group_colours) +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(cover_long$year)) +
   labs(y = "Percent Cover \n",
        x = "Year") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

@@

## Create Data Frames
```{r create data}
alpha_diversity_quad_macro <-
  data.frame(
    year = macroalgae$year,
    location = macroalgae$location,
    shannon = diversity(macroalgae[ , -which(names(macroalgae) %in% meta_cols)], "shannon"),
    richness = rowSums(macroalgae[ , -which(names(macroalgae) %in% meta_cols)] > 0)
  )

@@

## Figure 2a

@@

```{r richness df}
quad_richness_df <- alpha_diversity_quad_macro %>% 
  group_by(year, habitat) %>% 
  summarise(richness_mean = mean(richness),
            richness_se = se(richness))
```

### plot
```{r figure 2}
(figure_2a <- ggplot() +
  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 3) +
  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
                                             colour = habitat), width = 0, size = 1.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
   labs(y = "Macroalgal Richness \n",
        x = "Year",
        title = "a. Quadrat-level Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

## Figure 2b
### prep

@@

@@

## set up data for model
```{r quad data setup}
macroalgal_cover <- cover_df[,c("year", "habitat", "site", "location", "macroalgae")]

>>>>

## plot
```{r figure 1, fig.height = 10, fig.width=14, warning=FALSE}
(figure_1 <- ggplot() +
  geom_point(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
  geom_errorbar(data = figure_1_data, aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group),
                width = 0, size = 1.3) +
   geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 1) +
  facet_wrap(~ habitat) +
  annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
  scale_colour_manual(values = functional_group_colours) +
  # scale_fill_manual(values = functional_group_colours) +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(cover_long$year)) +
   labs(y = "Percent Cover \n",
        x = "Year") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

@@

## Create Data Frames
```{r create data for figure 2 quad}
alpha_diversity_quad_macro <-
  data.frame(
    year = macroalgae$year,
    location = macroalgae$location,
    shannon = diversity(macroalgae[, -which(names(macroalgae) %in% meta_cols)], "shannon"),
    richness = rowSums(macroalgae[, -which(names(macroalgae) %in% meta_cols)] > 0)
  )

#JG Add Evenness
alpha_diversity_quad_macro$ev <- alpha_diversity_quad_macro$shannon /
  (log(alpha_diversity_quad_macro$richness))

#makes 1 species 0 and 0 species NA for evenness metric
alpha_diversity_quad_macro$evenness <- ifelse(
  alpha_diversity_quad_macro$richness == 0,
  NA,
  # If richness == 0, ev is NA
  ifelse(
    alpha_diversity_quad_macro$richness == 1,
    0,
    # If richness == 1, ev is 0
    alpha_diversity_quad_macro$ev
  )
)

@@

```{r site level evenness}
#JG Add Evenness
alpha_diversity_site_macro$ev <- alpha_diversity_site_macro$shannon /
  (log(alpha_diversity_site_macro$richness))

#makes 1 species 0 and 0 species NA for evenness metric
alpha_diversity_site_macro$evenness <- ifelse(
  alpha_diversity_site_macro$richness == 0,
  NA,
  # If richness == 0, ev is NA
  ifelse(
    alpha_diversity_site_macro$richness == 1,
    0,
    # If richness == 1, ev is 0
    alpha_diversity_site_macro$ev
  )
)

alpha_diversity_site_macro <- left_join(site_meta_df, alpha_diversity_site_macro,by = join_by(year, site, habitat, site_habitat))
```

```{r}
macroalgal_cover <- cover_df[,c("year", "habitat", "site", "location", "macroalgae")]

## Figure 2a Richness @ Quad by Year

### prep
```{r richness df}
quad_richness_df <- alpha_diversity_quad_macro %>% 
  group_by(year, habitat) %>% 
  summarise(richness_mean = mean(richness),
            richness_se = se(richness))
```

### plot
```{r figure 2}
(figure_2a <- ggplot() +
  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 3) +
  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
                                             colour = habitat), width = 0, size = 1.3) +
  scale_colour_manual(values = habitat_colours) +
  # scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
   labs(y = "Macroalgal Richness \n",
        x = "Year",
        title = "a. Quadrat-level Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

## Figure 2b Richness @ Site by Year

@@

@@

## 2c Evenness @ Quad by Year

```{r calculate evennes}
#dear callie, why is this here?
site_richness_df <- alpha_diversity_site_macro %>% 
  group_by(year, habitat) %>% 
  summarise(richness_mean = mean(richness),
            richness_se = se(richness))

evenness_df <- alpha_diversity_quad_macro_with_cover %>% 
  drop_na(evenness) %>%
  group_by(year, habitat) %>% 
  summarise(eveness_mean = mean(evenness),
            eveness_se = se(evenness))

evenness_df$habitat <- factor(evenness_df$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

```{r plot evenness}
(figure_2c <- ggplot() +
  geom_line(data = evenness_df, aes(x = year, y = eveness_mean, colour = habitat), linewidth = 1) +
  geom_ribbon(data = evenness_df, aes(x = year, ymin = eveness_mean - eveness_se, ymax = eveness_mean + eveness_se, fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 0.7, yend = 0.7, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(evenness_df$year)) +
   labs(y = "Macroalgal Evenness \n",
        x = "Year") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

#ggsave(filename = "output/figure_2c.png", figure_2c, height = 10, width = 14)

#hist(eveness_df$eveness_mean)
```

## 2d Evenness @ Site by Year

```{r COME BACK HERE set up evenness}
site_evenness_df <- alpha_diversity_site_macro %>% 
  drop_na(evenness) %>%
  group_by(year, habitat) %>% 
  summarise(eveness_mean = mean(evenness),
            eveness_se = se(evenness))

site_evenness_df$habitat <- factor(site_evenness_df$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

```{r plot evenness}
(figure_2d <- ggplot() +
  geom_line(data = site_evenness_df, aes(x = year, y = eveness_mean, colour = habitat), linewidth = 1) +
  geom_ribbon(data = site_evenness_df, aes(x = year, ymin = eveness_mean - eveness_se, ymax = eveness_mean + eveness_se, fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 1, yend = 1, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(evenness_df$year)) +
   labs(y = "Macroalgal Evenness \n",
        x = "Year") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

#ggsave(filename = "output/figure_2d.png", figure_2d, height = 10, width = 14)
```

@@

## set up data for model
